---
title: "DEG_analysis"
author: "Tyler Audet"
date: "2024-11-14"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

```{r}
library(tidyverse)
library(data.table)
library(cowplot)
library(ashr)
library(patchwork)
library(glmmTMB)
library(emmeans)
```

First load and clean the M-F contrasts

!! Do not run, load in estimates below with ashr estimates !!
```{r}
dp <- na.omit(fread("../../outputs/pro_mf_contrast.txt"))
dm <- na.omit(fread("../../outputs/mel_mf_contrast.txt"))
dc <- na.omit(fread("../../outputs/car_mf_contrast.txt"))

dp <- dp[,c(2,6,7,8,9,11,12,13)]
dm <- dm[,c(2,6,7,8,9,11,12,13)]
dc <- dc[,c(2,6,7,8,9,11,12,13)]

headers <- c("gene","leg","stage","estimate","SE","LCL","UCL","p")

colnames(dp) <- headers
colnames(dm) <- headers
colnames(dc) <- headers

# Need to shrink, then figure out the UCL and LCL
dp <- cbind(dp, ash(betahat = dp$estimate, sebetahat = dp$SE, method = "shrink")$result)
dm <- cbind(dm, ash(betahat = dm$estimate, sebetahat = dm$SE, method = "shrink")$result)
dc <- cbind(dc, ash(betahat = dc$estimate, sebetahat = dc$SE, method = "shrink")$result)

dp <- cbind(dp, ashci(ash(betahat = dp$estimate, sebetahat = dp$SE, method = "shrink")))
dm <- cbind(dm, ashci(ash(betahat = dm$estimate, sebetahat = dm$SE, method = "shrink")))
dc <- cbind(dc, ashci(ash(betahat = dc$estimate, sebetahat = dc$SE, method = "shrink")))

dp$estimate_log2 <- log2(exp(dp$PosteriorMean))
dm$estimate_log2 <- log2(exp(dm$PosteriorMean))
dc$estimate_log2 <- log2(exp(dc$PosteriorMean))

dp$UCL_log2 <- log2(exp(dp$"0.975"))
dm$UCL_log2 <- log2(exp(dm$"0.975"))
dc$UCL_log2 <- log2(exp(dc$"0.975"))

dp$LCL_log2 <- log2(exp(dp$"0.025"))
dm$LCL_log2 <- log2(exp(dm$"0.025"))
dc$LCL_log2 <- log2(exp(dc$"0.025"))

write.csv(dp, "../../data/prolongata_norms")
write.csv(dc, "../../data/carrolli_norms")
write.csv(dm, "../../data/melanogaster_norms")
```


```{r}
dp <- fread("../../data/prolongata_norms")
dc <- fread("../../data/carrolli_norms")
dm <- fread("../../data/melanogaster_norms")
```

I am going to look at the distribution of these estimates
```{r}

hist(dp[leg == "for" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dp[leg == "for" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dp[leg == "mid" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dp[leg == "mid" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))

hist(dc[leg == "for" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dc[leg == "for" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dc[leg == "mid" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dc[leg == "mid" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))

hist(dm[leg == "for" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dm[leg == "for" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dm[leg == "mid" & stage == "late"]$estimate_log2, breaks = 500, xlim = c(-2,2))
hist(dm[leg == "mid" & stage == "early"]$estimate_log2, breaks = 500, xlim = c(-2,2))

```

Next load and clean the emmeans estimates without any contrasts
```{r}
mf_dp <- na.omit(fread("../../outputs/pro_separate.txt"))
mf_dm <- na.omit(fread("../../outputs/mel_separate.txt"))
mf_dc <- na.omit(fread("../../outputs/car_separate.txt"))

mf_dp <- mf_dp[,c(-1,-7,-8)]
mf_dm <- mf_dm[,c(-1,-7,-8)]
mf_dc <- mf_dc[,c(-1,-7,-8)]

colnames(mf_dp) <- c("gene", "sex", "leg", "stage", "estimate", "LCL", "UCL")
colnames(mf_dm) <- c("gene", "sex", "leg", "stage", "estimate", "LCL", "UCL")
colnames(mf_dc) <- c("gene", "sex", "leg", "stage", "estimate", "LCL", "UCL")

mf_dp$estimate_log2 <- log2(exp(mf_dp$estimate))
mf_dm$estimate_log2 <- log2(exp(mf_dm$estimate))
mf_dc$estimate_log2 <- log2(exp(mf_dc$estimate))

mf_dp$UCL_log2 <- log2(exp(mf_dp$UCL))
mf_dm$UCL_log2 <- log2(exp(mf_dm$UCL))
mf_dc$UCL_log2 <- log2(exp(mf_dc$UCL))

mf_dp$LCL_log2 <- log2(exp(mf_dp$LCL))
mf_dm$LCL_log2 <- log2(exp(mf_dm$LCL))
mf_dc$LCL_log2 <- log2(exp(mf_dc$LCL))
```

```{r}

# How many genes do we have values for and how many are male or female biased
length(unique(dp$gene)) # 10935 gff #10933 gffread #10952 agat
length(unique(dp$gene[dp$LCL_log2 > 0])) # 2845 MB
length(unique(dp$gene[dp$LCL_log2 > 0 & dp$leg == "for"])) # 2248
length(unique(dp$gene[dp$UCL_log2 < 0])) # 2820 FB
length(unique(dp$gene[dp$UCL_log2 < 0 & dp$leg == "for"])) # 1920

length(unique(dm$gene)) # 10768
length(unique(dm$gene[dm$LCL_log2 > 0])) # 1412 MB
length(unique(dm$gene[dm$LCL_log2 > 0 & dm$leg == "for"])) # 1123
length(unique(dm$gene[dm$UCL_log2 < 0])) # 1559 FB
length(unique(dm$gene[dm$UCL_log2 < 0 & dm$leg == "for"])) # 1189

length(unique(dc$gene)) # 6177 gff #10077 gffread gtf
length(unique(dc$gene[dc$LCL_log2 > 0])) # 1745 MB
length(unique(dc$gene[dc$LCL_log2 > 0 & dc$leg == "for"])) # 1590
length(unique(dc$gene[dc$UCL_log2 < 0])) # 1062 FB
length(unique(dc$gene[dc$UCL_log2 < 0 & dc$leg == "for"])) # 930

#doing the same thing for just genes that doubled their expression

dp_dbl <- dp[dp$UCL_log2 > 1 & dp$LCL_log2 > 1 | dp$LCL_log2 < -1 & dp$UCL_log2 < -1 & dp$leg == "for"]
length(unique(dp_dbl$gene)) # 631 gff # 632 gffread gtf #283 shrunk
dm_dbl <- dm[dm$UCL_log2 > 1 & dm$LCL_log2 > 1 | dm$LCL_log2 < -1 & dm$UCL_log2 < -1]
length(unique(dm_dbl$gene)) # 341 gff #60 shrunk?
dc_dbl <- dc[dc$UCL_log2 > 1 & dc$LCL_log2 > 1 | dc$LCL_log2 < -1 & dc$UCL_log2 < -1]
length(unique(dc_dbl$gene)) # 202 gff #388 gffread gtf #52 shrunk

# pulling out genes of interest with log2 fold change
dp_no_dc <- dp_dbl[-which(dp_dbl$gene %in% dc_dbl$gene)]
dp_interestingGenes <- dp_no_dc[-which(dp_no_dc$gene %in% dm_dbl$gene)]
dp_interestingGenes_for <- dp_interestingGenes[leg == "for"]
dp_interestingGenes_mid <- dp_interestingGenes[leg == "mid"]
dp_interestingGenes <- dp_interestingGenes_for[-which(dp_interestingGenes_for$gene %in% dp_interestingGenes_mid$gene)]
write.csv(dp_interestingGenes,file = "../outputs/log2DiffExp_prolongata_genes.csv",quote = F,row.names = F)

dp_interestingGenes_for$gene[which(dp_interestingGenes_for$gene %in% dp_interestingGenes_mid$gene)]

# sex specific doubling
male_biased_dp_x2 <- dp_dbl[dp_dbl$LCL_log2 > 1]
length(unique(male_biased_dp_x2$gene)) # 173
male_biased_dp_x2_for <- male_biased_dp_x2[leg == "for"]
male_biased_dp_x2_mid <- male_biased_dp_x2[leg == "mid"]
length(male_biased_dp_x2_for$gene[-which(male_biased_dp_x2_for$gene %in% male_biased_dp_x2_mid$gene)])

female_biased_dp_x2 <- dp_dbl[dp_dbl$UCL_log2 < -1]
length(unique(female_biased_dp_x2$gene)) # 380 female biased # # no outliers = 381
female_biased_dp_x2_for <- female_biased_dp_x2[leg == "for"]
female_biased_dp_x2_mid <- female_biased_dp_x2[leg == "mid"]
#length(female_biased_dp_x2_for$gene[-which(female_biased_dp_x2_for$gene %in% female_biased_dp_x2_mid$gene)])

male_biased_dc_x2 <- dc_dbl[dc_dbl$LCL_log2 > 1]
length(unique(male_biased_dc_x2$gene)) # 173
male_biased_dc_x2_for <- male_biased_dc_x2[leg == "for"]
male_biased_dc_x2_mid <- male_biased_dc_x2[leg == "mid"]
length(male_biased_dc_x2_for$gene[-which(male_biased_dc_x2_for$gene %in% male_biased_dc_x2_mid$gene)])

female_biased_dc_x2 <- dc_dbl[dc_dbl$UCL_log2 < -1]
length(unique(female_biased_dc_x2$gene)) # 380 female biased # # no outliers = 381
female_biased_dc_x2_for <- female_biased_dc_x2[leg == "for"]
female_biased_dc_x2_mid <- female_biased_dc_x2[leg == "mid"]
length(female_biased_dc_x2_for$gene[-which(female_biased_dc_x2_for$gene %in% female_biased_dc_x2_mid$gene)])

#

male_biased_dm_x2 <- dm_dbl[dm_dbl$LCL_log2 > 1]
length(unique(male_biased_dm_x2$gene)) # 173
male_biased_dm_x2_for <- male_biased_dm_x2[leg == "for"]
male_biased_dm_x2_mid <- male_biased_dm_x2[leg == "mid"]
length(male_biased_dm_x2_for$gene[-which(male_biased_dm_x2_for$gene %in% male_biased_dm_x2_mid$gene)])

female_biased_dm_x2 <- dm_dbl[dm_dbl$UCL_log2 < -1]
length(unique(female_biased_dm_x2$gene)) # 380 female biased # # no outliers = 381
female_biased_dm_x2_for <- female_biased_dm_x2[leg == "for"]
female_biased_dm_x2_mid <- female_biased_dm_x2[leg == "mid"]
length(female_biased_dm_x2_for$gene[-which(female_biased_dm_x2_for$gene %in% female_biased_dm_x2_mid$gene)])


genes_of_interest_male <- male_biased_dp_x2_for$gene[-which(male_biased_dp_x2_for$gene %in% male_biased_dp_x2_mid$gene)]

write.table(genes_of_interest_male, "../../outputs/genes_of_interest_male", quote = F, col.names = F, row.names = F)

genes_of_interest_female <- female_biased_dp_x2_for$gene

write.table(genes_of_interest_female, "../../outputs/genes_of_interest_female", quote = F, col.names = F, row.names = F)

#
```

I want to find the magnitude of expression for each sex in each leg, stage, and species

```{r}

# first I need all biased genes broken down into each sex, stage, and leg

dp_mb_for_early <- dp$estimate_log2[dp$LCL_log2 > 1 & dp$leg == "for" & dp$stage == "early"]
dp_mb_mid_early <- dp$estimate_log2[dp$LCL_log2 > 1 & dp$leg == "mid" & dp$stage == "early"]
dp_mb_for_late <- dp$estimate_log2[dp$LCL_log2 > 1 & dp$leg == "for" & dp$stage == "late"]
dp_mb_mid_late <- dp$estimate_log2[dp$LCL_log2 > 1 & dp$leg == "mid" & dp$stage == "late"]

dp_fb_for_early <- dp$estimate_log2[dp$UCL_log2 < -1 & dp$leg == "for" & dp$stage == "early"]
dp_fb_mid_early <- dp$estimate_log2[dp$UCL_log2 < -1 & dp$leg == "mid" & dp$stage == "early"]
dp_fb_for_late <- dp$estimate_log2[dp$UCL_log2 < -1 & dp$leg == "for" & dp$stage == "late"]
dp_fb_mid_late <- dp$estimate_log2[dp$UCL_log2 < -1 & dp$leg == "mid" & dp$stage == "late"]

dc_mb_for_early <- dc$estimate_log2[dc$LCL_log2 > 1 & dc$leg == "for" & dc$stage == "early"]
dc_mb_mid_early <- dc$estimate_log2[dc$LCL_log2 > 1 & dc$leg == "mid" & dc$stage == "early"]
dc_mb_for_late <- dc$estimate_log2[dc$LCL_log2 > 1 & dc$leg == "for" & dc$stage == "late"]
dc_mb_mid_late <- dc$estimate_log2[dc$LCL_log2 > 1 & dc$leg == "mid" & dc$stage == "late"]

dc_fb_for_early <- dc$estimate_log2[dc$UCL_log2 < -1 & dc$leg == "for" & dc$stage == "early"]
dc_fb_mid_early <- dc$estimate_log2[dc$UCL_log2 < -1 & dc$leg == "mid" & dc$stage == "early"]
dc_fb_for_late <- dc$estimate_log2[dc$UCL_log2 < -1 & dc$leg == "for" & dc$stage == "late"]
dc_fb_mid_late <- dc$estimate_log2[dc$UCL_log2 < -1 & dc$leg == "mid" & dc$stage == "late"]

dm_mb_for_early <- dm$estimate_log2[dm$LCL_log2 > 1 & dm$leg == "for" & dm$stage == "early"]
dm_mb_mid_early <- dm$estimate_log2[dm$LCL_log2 > 1 & dm$leg == "mid" & dm$stage == "early"]
dm_mb_for_late <- dm$estimate_log2[dm$LCL_log2 > 1 & dm$leg == "for" & dm$stage == "late"]
dm_mb_mid_late <- dm$estimate_log2[dm$LCL_log2 > 1 & dm$leg == "mid" & dm$stage == "late"]

dm_fb_for_early <- dm$estimate_log2[dm$UCL_log2 < -1 & dm$leg == "for" & dm$stage == "early"]
dm_fb_mid_early <- dm$estimate_log2[dm$UCL_log2 < -1 & dm$leg == "mid" & dm$stage == "early"]
dm_fb_for_late <- dm$estimate_log2[dm$UCL_log2 < -1 & dm$leg == "for" & dm$stage == "late"]
dm_fb_mid_late <- dm$estimate_log2[dm$UCL_log2 < -1 & dm$leg == "mid" & dm$stage == "late"]

templist <- list(dp_mb_for_early,dp_mb_mid_early,dp_mb_for_late,dp_mb_mid_late,dp_fb_for_early,dp_fb_mid_early,dp_fb_for_late,dp_fb_mid_late,dc_mb_for_early,dc_mb_mid_early,dc_mb_for_late,dc_mb_mid_late,dc_fb_for_early,dc_fb_mid_early,dc_fb_for_late,dc_fb_mid_late,dm_mb_for_early,dm_mb_mid_early,dm_mb_for_late,dm_mb_mid_late,dm_fb_for_early,dm_fb_mid_early,dm_fb_for_late,dm_fb_mid_late)

easy_norm <- function(x) {sqrt(sum(x^2))/length(x)}

#lapply(templist, easy_norm)

# sampling with replacement for bootstrapping 
boot_norm_function <- function(x){
    x_resample <- sample(x, replace = TRUE)
    norm_resample <- norm(as.matrix(x_resample), type = "2")/length(x_resample)
    return(norm_resample)
  }

mag_mat <- matrix(nrow = 24, ncol = 3)
x <- 1

for (i in templist) {
  mag_mat[x,1] <- easy_norm(i)
  mag_mat[x,2:3] <- quantile(replicate(10000, boot_norm_function(i)), probs = c(0.025, 0.975), na.rm = T)
  x <- x + 1
}

row.names(mag_mat) <- c("dp_mb_for_early","dp_mb_mid_early","dp_mb_for_late","dp_mb_mid_late","dp_fb_for_early","dp_fb_mid_early","dp_fb_for_late","dp_fb_mid_late","dc_mb_for_early","dc_mb_mid_early","dc_mb_for_late","dc_mb_mid_late","dc_fb_for_early","dc_fb_mid_early","dc_fb_for_late","dc_fb_mid_late","dm_mb_for_early","dm_mb_mid_early","dm_mb_for_late","dm_mb_mid_late","dm_fb_for_early","dm_fb_mid_early","dm_fb_for_late","dm_fb_mid_late")

colnames(mag_mat) <- c("magnitude","LCL","UCL")

SSD <- fread("../../data/rhop_clade_SSD.csv")

SSD <- SSD[cohort == "C0" & trait != "thoraxL" & species != "fuy"]

pro_fore <- sum(SSD$SSD_measure[SSD$species =="pro"][1:2]) * -1
pro_mid <- sum(SSD$SSD_measure[SSD$species =="pro"][7:8]) * -1

car_fore <- sum(SSD$SSD_measure[SSD$species =="car"][1:2]) * -1
car_mid <- sum(SSD$SSD_measure[SSD$species =="car"][7:8]) * -1

mel_SSD <- fread("../../data/pengal4_clean.csv")

mel_SSD <- mel_SSD[code == "control"]

mel_SSD$sex <- relevel(as.factor(mel_SSD$sex), ref = "M")

mel_mod <- glmmTMB(length ~ trait*sex + (1|rep),
        data = mel_SSD)

mel_contrast <- as.data.table(emmeans(mel_mod, pairwise ~ sex | trait)[[2]])

mel_fore <- sum(mel_contrast$estimate[1:2])
mel_mid <- sum(mel_contrast$estimate[3:4])

SSD_estimates <- c(rep(c(pro_fore,pro_mid), 4),rep(c(car_fore,car_mid), 4),rep(c(mel_fore,mel_mid), 4))

sexes <- rep(c(rep("male",4), rep("female", 4)), 3)
SBGs <- unlist(lapply(list(dp_mb_for_early,dp_mb_mid_early,dp_mb_for_late,dp_mb_mid_late,dp_fb_for_early,dp_fb_mid_early,dp_fb_for_late,dp_fb_mid_late,dc_mb_for_early,dc_mb_mid_early,dc_mb_for_late,dc_mb_mid_late,dc_fb_for_early,dc_fb_mid_early,dc_fb_for_late,dc_fb_mid_late,dm_mb_for_early,dm_mb_mid_early,dm_mb_for_late,dm_mb_mid_late,dm_fb_for_early,dm_fb_mid_early,dm_fb_for_late,dm_fb_mid_late), length))

mag_mat_final <- as.data.table(cbind(mag_mat,SSD_estimates,c(rep("pro",8),rep("car",8),rep("mel",8)), sexes, rep(c("fore","mid"),12), rep(c("early","early","late","late"), 6), SBGs))


mag_mat_final[,1:4] <- lapply(mag_mat_final[,1:4], as.numeric)
colnames(mag_mat_final)[5] <- "species"
colnames(mag_mat_final)[6] <- "sex"
colnames(mag_mat_final)[7] <- "leg"
colnames(mag_mat_final)[8] <- "stage"
colnames(mag_mat_final)[9] <- "SBGs"

colour_vector <- c("fore.female" = "#E66100",
                   "mid.female" = "orange",
                   "fore.male" = "#5D3A9B",
                   "mid.male" = "orchid")

shape_vector <- c("car.early" = 15,
                  "mel.early" = 17,
                  "pro.early" = 16,
                  "car.late" = 0,
                  "mel.late" = 2,
                  "pro.late" = 1)


mag_plot <- ggplot(mag_mat_final, aes(x = SSD_estimates, y = magnitude, colour = interaction(leg,sex), shape = interaction(species, stage, drop = TRUE) )) +
  geom_pointrange(aes(ymin = LCL, ymax = UCL), position = position_dodge2(width = 0.075), size = 0.75, alpha = 0.7) +
  ylab("Magnitude of SBGE") +
  xlab("Femur SSD (M-F)") +
  scale_colour_manual(values = colour_vector, labels = c("Female foreleg","Female midleg","Male foreleg","Male midleg"))  +
  scale_shape_manual(values = shape_vector, labels = c(expression(paste(italic("D. carrolli"), " early")), expression(paste(italic("D. melanogaster"), " early")), expression(paste(italic("D. prolongata"), " early")),
expression(paste(italic("D. carrolli"), " late")), expression(paste(italic("D. melanogaster"), " late")), expression(paste(italic("D. prolongata"), " late"))), name = "Species and stage") +
  theme_classic() +
  labs(shape="Species and stage", colour="Sex and leg")

SBG_plot <- ggplot(mag_mat_final, aes(x = SSD_estimates, y = as.numeric(SBGs), colour = interaction(leg,sex), shape = interaction(species, stage, drop = TRUE) )) +
  geom_pointrange(aes(ymin = LCL, ymax = UCL), position = position_dodge2(width = 0.075), size = 0.75, alpha = 0.7) +
  ylab("Number of SBGs") +
  xlab("Femur SSD (M-F)") +
  scale_colour_manual(values = colour_vector, labels = c("Female foreleg","Female midleg","Male foreleg","Male midleg"))  +
  scale_shape_manual(values = shape_vector, labels = c(expression(paste(italic("D. carrolli"), " early")), expression(paste(italic("D. melanogaster"), " early")), expression(paste(italic("D. prolongata"), " early")),
expression(paste(italic("D. carrolli"), " late")), expression(paste(italic("D. melanogaster"), " late")), expression(paste(italic("D. prolongata"), " late"))), name = "Species and stage") +
  theme_classic() +
  labs(shape="Species and stage", colour="Sex and leg")

SBG_plot+mag_plot + plot_layout(guides = 'collect')

```


# plotting reaction norms for L2F D. prolongata genes compared to other species
```{r}
#log2_genes <- fread("../../outputs/log2DiffExp_prolongata_genes.csv")

genesOfInterest <- unique(dp_interestingGenes$gene)


dp_complete <- dp[which(dp$gene %in% genesOfInterest)]
dp_complete$species <- "prolongata"
dm_complete <- dm[which(dm$gene %in% genesOfInterest)]
dm_complete$species <- "melanogaster"
dc_complete <- dc[which(dc$gene %in% genesOfInterest)]
dc_complete$species <- "carrolli"

allRows <- rbind(dp_complete,dm_complete,dc_complete)

step1 <- melt(allRows, id.vars = c("gene","leg","stage","species"), measure.vars = c("estimate_log2","LCL_log2","UCL_log2","p"))

plotting_data <- dcast(step1, gene + stage + species + leg ~ variable, mean)

colnames(plotting_data) <- c("gene","stage","species","leg","estimate","LCL","UCL","p")

pdf("../../paper/Final_Figures/Supplemental_reactionNorms.pdf",width=5,height=3)

for(i in unique(plotting_data$gene)) {

this_gene <- i

current_gene <- plotting_data[plotting_data$gene == this_gene,] 


geneplot <- ggplot(current_gene, aes(x = leg, y = estimate)) +
  geom_pointrange(aes(color = species,ymin = LCL, ymax = UCL), size = 0.2, position = position_dodge(width = 0.5)) +
  labs(title = this_gene) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species, colour = species), fun = mean, geom = "path",position = position_dodge(width = 0.5))

print(geneplot)

}

dev.off()
```

Below I plot a subset of genes that we tested RNAi lines for
```{r}

figure_genes <- c("dysf","grn","Sox15","CG30457","CG13285","bab1","otp")

dp_complete <- dp[which(dp$gene %in% figure_genes)]
dp_complete$species <- "prolongata"
dm_complete <- dm[which(dm$gene %in% figure_genes)]
dm_complete$species <- "melanogaster"
dc_complete <- dc[which(dc$gene %in% figure_genes)]
dc_complete$species <- "carrolli"

allRows <- rbind(dp_complete,dm_complete,dc_complete)

step1 <- melt(allRows, id.vars = c("gene","leg","stage","species"), measure.vars = c("estimate_log2","LCL_log2","UCL_log2","p"))

plotting_data <- dcast(step1, gene + stage + species + leg ~ variable, mean)

colnames(plotting_data) <- c("gene","stage","species","leg","estimate","LCL","UCL","p")

plotting_data$species <- gsub('^', 'D. ', plotting_data$species)

#plotting_data$species <- expression(italic(plotting_data$species))

plotting_data$leg <- sub("for","foreleg",plotting_data$leg)
plotting_data$leg <- sub("mid","midleg",plotting_data$leg)

otp <- filter(plotting_data, gene == "otp") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species, ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("orthopedia"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
 theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  ylab(NULL) +
  xlab(NULL)

dysf <- filter(plotting_data, gene == "dysf") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("dysfusion"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
 theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  ylab(NULL) +
  xlab(NULL)

grn <- filter(plotting_data, gene == "grn") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("grain"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
 theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  ylab(NULL) +
  xlab(NULL)

Sox15 <- filter(plotting_data, gene == "Sox15") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("sox box protein 15"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  ylab(NULL) +
  xlab(NULL)

CG30457 <- filter(plotting_data, gene == "CG30457") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("CG30457"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
 theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  xlab(NULL) +
  ylab("Gene expression estimate\n(Male-Female)")

CG13285 <- filter(plotting_data, gene == "CG13285") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("CG13285"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  ylab(NULL) +
  xlab(NULL)

bab1 <- filter(plotting_data, gene == "bab1") %>%
ggplot(aes(x = leg, y = estimate)) +
  geom_pointrange(aes(shape = species,ymin = LCL, ymax = UCL), size = 0.4, position = position_dodge(width = 0.5)) +
  labs(title = expression(italic("bric-a-brac1"))) +
  facet_wrap(~stage) +
  stat_summary(aes(group = species), fun = mean, geom = "path",position = position_dodge(width = 0.5)) +
  theme_classic() +
  scale_shape_manual(values = c(15,17,16), name = "Species") +
  theme(legend.text = element_text(face = "italic", size = 15), legend.title = element_text(size = 15)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  ylab(NULL) +
  xlab(NULL)

design <- "ab
           cd
           ef
           gh"

dysf+
  grn+
  CG30457+
  CG13285+
  Sox15+
  bab1+
  otp+
  guide_area()+
  plot_layout(
    design = design, 
    guides = "collect"
  )


```

The below block of code plots reaction norms for the sexes separately to look at expression levels rather than a contrast.

```{r}

mf_dp_complete <- mf_dp[which(mf_dp$gene %in% genesOfInterest)]
mf_dp_complete$species <- "prolongata"
mf_dm_complete <- mf_dm[which(mf_dm$gene %in% genesOfInterest)]
mf_dm_complete$species <- "melanogaster"
mf_dc_complete <- mf_dc[which(mf_dc$gene %in% genesOfInterest)]
mf_dc_complete$species <- "carrolli"



mf_allRows <- rbind(mf_dp_complete,mf_dm_complete,mf_dc_complete)

step1 <- melt(mf_allRows, id.vars = c("gene","sex","leg","stage","species"), measure.vars = c("estimate_log2","LCL_log2","UCL_log2"))


plotting_data <- dcast(step1, gene + sex + stage + species + leg ~ variable, sum)

pdf("../../paper/Figures/DEG_reactionNorms_MFseperate.pdf",width=5,height=3)

for(i in unique(plotting_data$gene)) {

this_gene <- i

current_gene <- plotting_data[plotting_data$gene == this_gene,] 


geneplot <- ggplot(current_gene, aes(x = leg, y = estimate_log2, shape = sex)) +
  geom_pointrange(aes(color = species, ymin = LCL_log2, ymax = UCL_log2), size = 0.2, position = position_dodge(width = 0.5)) +
  labs(title = this_gene) +
  facet_wrap(~stage) +
  stat_summary(aes(group = interaction(species, sex), colour = species), fun = mean, geom = "line", position = position_dodge2(width = 0.5))

print(geneplot)

}

dev.off()

```

Below is the scripts for the MA plots with ashr estimates

```{r}
# MA plotting

temp_dp <- dcast(mf_dp, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dp$mean <- (temp_dp$F + temp_dp$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dp <- merge(temp_dp, dp, by = c("gene","leg","stage"))

dp_FE <-MA_dp %>%
        filter(leg == "for" & stage == "early") %>%
        ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
          ylim(c(-10, 10)) +
          xlim(c(0,25)) +
          ylab("Log2 M-F prolongata") +
          xlab(NULL) +
          scale_color_manual(values=c("darkgrey","darkred")) +
          theme_classic() +
          geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_FL <-MA_dp %>%
  filter(leg == "for" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_ME <-MA_dp %>%
  filter(leg == "mid" & stage == "early") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F prolongata") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_ML <-MA_dp %>%
  filter(leg == "mid" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)


temp_dm <- dcast(mf_dm, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dm$mean <- (temp_dm$F + temp_dm$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dm <- merge(temp_dm, dm, by = c("gene","leg","stage"))

dm_FE <-MA_dm %>%
  filter(leg == "for" & stage == "early") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("Log2 M-F melanogaster") +
  xlab("Log2 mean expression early") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_FL <-MA_dm %>%
  filter(leg == "for" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab("Log2 mean expression late") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_ME <-MA_dm %>%
  filter(leg == "mid" & stage == "early") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F melanogaster") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_ML <-MA_dm %>%
  filter(leg == "mid" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)



temp_dc <- dcast(mf_dc, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dc$mean <- (temp_dc$F + temp_dc$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dc <- merge(temp_dc, dc, by = c("gene","leg","stage"))


dc_FE <-MA_dc %>%
  filter(leg == "for" & stage == "early") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("Log2 M-F carrolli") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_FL <-MA_dc %>%
  filter(leg == "for" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_ME <-MA_dc %>%
  filter(leg == "mid" & stage == "early") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F carrolli") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_ML <-MA_dc %>%
  filter(leg == "mid" & stage == "late") %>%
  ggplot(aes(x = mean, y = estimate_log2, colour = LCL_log2 > 0 | UCL_log2 < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

plot_grid(dp_FE,dp_FL,dc_FE,dc_FL,dm_FE,dm_FL,ncol = 2)

plot_grid(dp_ME,dp_ML,dc_ME,dc_ML,dm_ME,dm_ML,ncol = 2)

```

Below is the scripts for the MA plots without ashr estimates

```{r}
# MA plotting

temp_dp <- dcast(mf_dp, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dp$mean <- (temp_dp$F + temp_dp$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dp <- merge(temp_dp, dp, by = c("gene","leg","stage"))

dp_FE <-MA_dp %>%
        filter(leg == "for" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
          ylim(c(-10, 10)) +
          xlim(c(0,25)) +
          ylab("Log2 M-F prolongata") +
          xlab(NULL) +
          scale_color_manual(values=c("darkgrey","darkred")) +
          theme_classic() +
          geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_FL <-MA_dp %>%
  filter(leg == "for" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_ME <-MA_dp %>%
  filter(leg == "mid" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F prolongata") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dp_ML <-MA_dp %>%
  filter(leg == "mid" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)


temp_dm <- dcast(mf_dm, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dm$mean <- (temp_dm$F + temp_dm$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dm <- merge(temp_dm, dm, by = c("gene","leg","stage"))

dm_FE <-MA_dm %>%
  filter(leg == "for" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("Log2 M-F melanogaster") +
  xlab("Log2 mean expression early") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_FL <-MA_dm %>%
  filter(leg == "for" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab("Log2 mean expression late") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_ME <-MA_dm %>%
  filter(leg == "mid" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F melanogaster") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dm_ML <-MA_dm %>%
  filter(leg == "mid" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)



temp_dc <- dcast(mf_dc, gene + leg + stage ~ sex, value.var = "estimate_log2")
temp_dc$mean <- (temp_dc$F + temp_dc$M) / 2
#temp_dp$fold <- (temp_dp$M - temp_dp$F)
MA_dc <- merge(temp_dc, dc, by = c("gene","leg","stage"))


dc_FE <-MA_dc %>%
  filter(leg == "for" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("Log2 M-F carrolli") +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_FL <-MA_dc %>%
  filter(leg == "for" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab(NULL) +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_ME <-MA_dc %>%
  filter(leg == "mid" & stage == "early") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab("FC M-F carrolli") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

dc_ML <-MA_dc %>%
  filter(leg == "mid" & stage == "late") %>%
        ggplot(aes(x = mean, y = log2(exp(estimate)), colour = log2(exp(LCL)) > 0 | log2(exp(UCL)) < 0 )) +
  ylim(c(-10, 10)) +
  xlim(c(0,25)) +
  ylab(NULL) +
  xlab("Log2 mean expression late") +
  scale_color_manual(values=c("darkgrey","darkred")) +
  theme_classic() +
  geom_point(alpha = 0.5, size = 0.5, show.legend = FALSE)

plot_grid(dp_FE,dp_FL,dc_FE,dc_FL,dm_FE,dm_FL,ncol = 2)

plot_grid(dp_ME,dp_ML,dc_ME,dc_ML,dm_ME,dm_ML,ncol = 2)

```
