---
title: "DeSeq Analysis"
author: "Tyler Audet"
date: "`r format(Sys.time(),'%d %b %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

This is the script used to verify our methods against standard DeSeq methods for the paper 'Genomic architecture of the developing forelegs of Drosophila prolongata; an exaggerated weapon and ornament'. This script reads in raw counts generated from the script `STAR_[species].sh` to run the DeSeq analyses as well as our shrunked estimates as the files `[species]_norms` generated from the script `DGE_analysis.Rmd`. It generates the supplemental figures S4 and S5.

```{r}
library(DESeq2)
library(edgeR)
library(tidyverse)
library(data.table)
```


I load in carrolli and get a DESeq object

```{r}
car_counts <- read.table("../../outputs/car_counts.txt")

car_names <- as.data.frame(names(as.data.frame(car_counts)))

colnames(car_names) <- "sample"

car_design <- separate(car_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

car_design$replicate <- sub('ReadsPerGene.out.tab', "", car_design$replicate)


car_design <- car_design[,-c(1:2)]

# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(car_counts))

#Determine which genes have sufficiently large counts to be retained in a 
smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I want to filter by the DGE_object to get genes that pass filters
car_counts <- car_counts[row.names(DGE_object$counts),]

#Need to make my formula for DeSeq
car_design[,1:4] <- lapply(car_design[,1:4], as.factor)

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
car_dds <- DESeqDataSetFromMatrix(countData = car_counts,
                                  colData = car_design,
                                  design = ~ sex * leg * stage)

car_dds <- DESeq2::estimateSizeFactors(car_dds)

colData(car_dds)

car_dds <- DESeq(car_dds)

resultsNames(car_dds)

```

I load in prolongata and get a DESeq object

```{r}
pro_counts <- read.table("../../outputs/pro_counts.txt")

pro_names <- as.data.frame(names(as.data.frame(pro_counts)))

colnames(pro_names) <- "sample"

pro_design <- separate(pro_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

pro_design$replicate <- sub('ReadsPerGene.out.tab', "", pro_design$replicate)


pro_design <- pro_design[,-c(1:2)]

# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(pro_counts))

#Determine which genes have sufficiently large counts to be retained in a 
smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I want to filter by the DGE_object to get genes that pass filters
pro_counts <- pro_counts[row.names(DGE_object$counts),]

#Need to make my formula for DeSeq
pro_design[,1:4] <- lapply(pro_design[,1:4], as.factor)

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
pro_dds <- DESeqDataSetFromMatrix(countData = pro_counts,
                                  colData = pro_design,
                                  design = ~ sex * leg * stage)

pro_dds <- DESeq2::estimateSizeFactors(pro_dds)

colData(pro_dds)

pro_dds <- DESeq(pro_dds)

resultsNames(pro_dds)
```

I load in melanogaster to get a DESeq object

```{r}
mel_counts <- read.table("../../outputs/mel_counts.txt")

mel_names <- as.data.frame(names(as.data.frame(mel_counts)))

colnames(mel_names) <- "sample"

mel_design <- separate(mel_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

mel_design$replicate <- sub('ReadsPerGene.out.tab', "", mel_design$replicate)


mel_design <- mel_design[,-c(1:2)]

# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(mel_counts))

#Determine which genes have sufficiently large counts to be retained in a 
smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I want to filter by the DGE_object to get genes that pass filters
mel_counts <- mel_counts[row.names(DGE_object$counts),]

#Need to make my formula for DeSeq
mel_design[,1:4] <- lapply(mel_design[,1:4], as.factor)

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
mel_dds <- DESeqDataSetFromMatrix(countData = mel_counts,
                                  colData = mel_design,
                                  design = ~ sex * leg * stage)

mel_dds <- DESeq2::estimateSizeFactors(mel_dds)

colData(mel_dds)

mel_dds <- DESeq(mel_dds)

resultsNames(mel_dds)
```

```{r}

test_for_early <- as.data.frame(results(pro_dds, contrast=list("sex_M_vs_F")))

# uncorrected for comparison
#test_for_early <- as.data.frame(results(pro_dds, contrast=list("sex_M_vs_F")))

#test_for_early_LF2 <- test_for_early[which(test_for_early$log2FoldChange > 1 | test_for_early$log2FoldChange < -1 & test_for_early$padj < 0.05),]

test_for_early$gene <- row.names(test_for_early)

#

test_for_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.stagelate"))))

# uncorrected for comparison
#test_for_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.stagelate"))))


#test_for_late_LF2 <- test_for_late[which(test_for_late$log2FoldChange > 1 | test_for_late$log2FoldChange < -1 & test_for_late$padj < 0.05),]

test_for_late$gene <- row.names(test_for_late)

#

test_mid_early <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid"))))

# uncorrected for comparison
#test_mid_early <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid"))))


#test_mid_early_LF2 <- test_mid_early[which(test_mid_early$log2FoldChange > 1 | test_mid_early$log2FoldChange < -1 & test_mid_early$padj < 0.05),]

test_mid_early$gene <- row.names(test_mid_early)

#

test_mid_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid.stagelate"))))

# uncorrected for comparison
#test_mid_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid.stagelate"))))

#test_mid_late_LF2 <- test_mid_late[which(test_mid_late$log2FoldChange > 1 | test_mid_late$log2FoldChange < -1 & test_mid_late$padj < 0.05),]

test_mid_late$gene <- row.names(test_mid_late)

#


# these are emmeans estimates > 2 log2 fold change
genes <- fread("../data/prolongata_norms", h = T)

genes_for_late <- genes[leg == "for" & stage == "late"]
genes_for_early <- genes[leg == "for" & stage == "early"]
genes_mid_early <- genes[leg == "mid" & stage == "early"]
genes_mid_late <- genes[leg == "mid" & stage == "late"]

#genes_for_late$gene <- rownames(genes_for_late)
#genes_for_early$gene <- rownames(genes_for_early)
#genes_mid_late$gene <- rownames(genes_mid_late)
#genes_mid_early$gene <- rownames(genes_mid_early)

# merge them for plotting, this also pulls out the genes found in both data.tables
interesting_genes_for_late <- merge(genes_for_late, test_for_late, by="gene")

interesting_genes_for_early <- merge(genes_for_early, test_for_early, by="gene")

interesting_genes_mid_late <- merge(genes_mid_late, test_mid_late, by="gene")

interesting_genes_mid_early <- merge(genes_mid_early, test_mid_early, by="gene")
```

non-regularized
```{r}
ggplot(interesting_genes_for_late, aes(x = estimate, y = log2FoldChange, colour = LCL > 1 | UCL < -1)) +
  geom_point(alpha = 0.5, size = 1.25) +
  ylab("DESeq estimate without ashr regularization") +
  xlab("glmmTMB estimate without ashr regularization") +
  geom_linerange(aes(xmin = LCL, xmax = UCL), alpha = 0.4) +
  geom_linerange(aes(ymin = log2FoldChange-2*lfcSE, ymax = log2FoldChange+2*lfcSE), alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, colour = "red", linetype = 2) +
  theme_classic() +
  theme(legend.position="none") +
  scale_colour_manual(values = c("lightgrey","black"))
  
ggplot(interesting_genes_for_early, aes(x = estimate, y = log2FoldChange, colour = LCL > 1 | UCL < -1)) +
  geom_point(alpha = 0.5, size = 1.25) +
  ylab("DESeq estimate without ashr regularization") +
  xlab("glmmTMB estimate without ashr regularization") +
  geom_linerange(aes(xmin = LCL, xmax = UCL), alpha = 0.4) +
  geom_linerange(aes(ymin = log2FoldChange-2*lfcSE, ymax = log2FoldChange+2*lfcSE), alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, colour = "red", linetype = 2) +
  theme_classic() +
  theme(legend.position="none") +
  scale_colour_manual(values = c("lightgrey","black"))
  
interesting_genes_for_early[interesting_genes_for_early$log2FoldChange > 10]

interesting_genes_for_early[interesting_genes_for_early$log2FoldChange > 10]

```


regularize

```{r}

test_for_early <- as.data.frame(lfcShrink(pro_dds,res = results(pro_dds, contrast=list("sex_M_vs_F")), type = "ashr"))

# uncorrected for comparison
#test_for_early <- as.data.frame(results(pro_dds, contrast=list("sex_M_vs_F")))

#test_for_early_LF2 <- test_for_early[which(test_for_early$log2FoldChange > 1 | test_for_early$log2FoldChange < -1 & test_for_early$padj < 0.05),]

test_for_early$gene <- row.names(test_for_early)

#

test_for_late <- as.data.frame(lfcShrink(pro_dds,res = results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.stagelate"))), type = "ashr"))

# uncorrected for comparison
#test_for_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.stagelate"))))


#test_for_late_LF2 <- test_for_late[which(test_for_late$log2FoldChange > 1 | test_for_late$log2FoldChange < -1 & test_for_late$padj < 0.05),]

test_for_late$gene <- row.names(test_for_late)

#

test_mid_early <- as.data.frame(lfcShrink(pro_dds,res = results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid"))), type = "ashr"))

# uncorrected for comparison
#test_mid_early <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid"))))


#test_mid_early_LF2 <- test_mid_early[which(test_mid_early$log2FoldChange > 1 | test_mid_early$log2FoldChange < -1 & test_mid_early$padj < 0.05),]

test_mid_early$gene <- row.names(test_mid_early)

#

test_mid_late <- as.data.frame(lfcShrink(pro_dds,res = results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid.stagelate"))), type = "ashr"))

# uncorrected for comparison
#test_mid_late <- as.data.frame(results(pro_dds, contrast=list(c("sex_M_vs_F","sexM.legmid.stagelate"))))

#test_mid_late_LF2 <- test_mid_late[which(test_mid_late$log2FoldChange > 1 | test_mid_late$log2FoldChange < -1 & test_mid_late$padj < 0.05),]

test_mid_late$gene <- row.names(test_mid_late)

```



```{r}

ggplot(interesting_genes_for_late, aes(x = estimate_log2, y = log2FoldChange, colour = LCL_log2 > 1 | UCL_log2 < -1)) +
  geom_point(alpha = 0.5, size = 1.25) +
  ylab("DESeq estimate with ashr regularization") +
  xlab("glmmTMB estimate with ashr regularization") +
  geom_linerange(aes(xmin = LCL_log2, xmax = UCL_log2), alpha = 0.4) +
  geom_linerange(aes(ymin = log2FoldChange-2*lfcSE, ymax = log2FoldChange+2*lfcSE), alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, colour = "red", linetype = 2) +
  theme_classic() +
  theme(legend.position="none") +
  scale_colour_manual(values = c("lightgrey","black"))
  
ggplot(interesting_genes_for_early, aes(x = estimate_log2, y = log2FoldChange, colour = LCL_log2 > 1 | UCL_log2 < -1)) +
  geom_point(alpha = 0.5, size = 1.25) +
  ylab("DESeq estimate with ashr regularization") +
  xlab("glmmTMB estimate with ashr regularization") +
  geom_linerange(aes(xmin = LCL_log2, xmax = UCL_log2), alpha = 0.4) +
  geom_linerange(aes(ymin = log2FoldChange-2*lfcSE, ymax = log2FoldChange+2*lfcSE), alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, colour = "red", linetype = 2) +
  theme_classic() +
  theme(legend.position="none") +
  scale_colour_manual(values = c("lightgrey","black"))
  
interesting_genes_for_early[interesting_genes_for_early$log2FoldChange > 10]

```



