---
title: "Cell_number_supp"
output: 
  html_document: 
    toc: true
    number_sections: true
    keep_md: true
date: "2025-11-12"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)
```

This is the script for the reanalysis of cell numbers for the paper 'Genomic architecture of the developing forelegs of Drosophila prolongata; an exaggerated weapon and ornament'. This script will take cell number and size data from Luecke & Kopp 2019 available via Dryad at [https://datadryad.org/dataset/doi:10.25338/B8303J] and model it to demonstrate that cell number is not a likely confounding factor for our gene expression results. This script outputs supplemental figure S1.

```{r}
library(ggplot2)
library(glmmTMB)
library(emmeans)
library(data.table)
library(patchwork)
library(bbmle)
```

```{r}
ddat <- fread("../../data/Luecke_DrosophilaLegSize_Discs_LarvaData.csv")
```


```{r}
ddat$T1SIZE <- (ddat$T1aSIZE + ddat$T1bSIZE) / 2
ddat$T2SIZE <- (ddat$T2aSIZE + ddat$T2bSIZE) / 2

ddat$T1DAPI <- (ddat$T1aDAPI + ddat$T1bDAPI) / 2
ddat$T2DAPI <- (ddat$T2aDAPI + ddat$T2bDAPI) / 2
```


```{r}
carM <- ddat[which(ddat$species == "kb866" & ddat$sex == "m" & method == "AEL")]
carF <- ddat[which(ddat$species == "kb866" & ddat$sex == "f" & method == "AEL")]
proM <- ddat[which(ddat$species == "prol" & ddat$sex == "m" & method == "AEL")]
proF <- ddat[which(ddat$species == "prol" & ddat$sex == "f" & method == "AEL")]
```


```{r}
carF$percent.dev <- as.numeric(carF$timepoint)/6.7
carM$percent.dev <- as.numeric(carM$timepoint)/7
proF$percent.dev <- as.numeric(proF$timepoint)/7.2
proM$percent.dev <- as.numeric(proM$timepoint)/8
```


```{r}
AEL <- rbind(carF,carM,proF,proM)
```


```{r}
carM.L3 <- ddat[which(ddat$species == "kb866" & ddat$sex == "m" & method == "AL3")]
carF.L3 <- ddat[which(ddat$species == "kb866" & ddat$sex == "f" & method == "AL3")]
proM.L3 <- ddat[which(ddat$species == "prol" & ddat$sex == "m" & method == "AL3")]
proF.L3 <- ddat[which(ddat$species == "prol" & ddat$sex == "f" & method == "AL3")]
```


```{r}
carF.L3$percent.dev <- (as.numeric(carF.L3$timepoint)+(6.7*24)/2)/(6.7*24)
carM.L3$percent.dev <- (as.numeric(carM.L3$timepoint)+((7*24)/2))/(7*24)
proF.L3$percent.dev <- (as.numeric(proF.L3$timepoint)+((7.2*24)/2))/(7.2*24)
proM.L3$percent.dev <- (as.numeric(proM.L3$timepoint)+((8*24)/2))/(8*24)
AL3<-rbind(carF.L3,carM.L3,proF.L3,proM.L3)

# late timepoint is 6.7 days AEL -> 0.838% of larval development
# early timepoint is 6.25 AEL -> 0.781% of larval development

wL3 <- ddat[ddat$timepoint == "wL3"]
wL3$percent.dev <- as.numeric("1")

wpp <- ddat[ddat$timepoint == "wpp"]
wpp$percent.dev <- as.numeric("1.025")

all.time <- rbind(AEL,AL3,wL3,wpp)

all.time$species <- sub("kb866", "D. carrolli", all.time$species)
all.time$species <- sub("prol", "D. prolongata", all.time$species)
all.time$sex <- sub("m", "Male", all.time$sex)
all.time$sex <- sub("f", "Female", all.time$sex)
```


```{r}
disc.size <- ggplot(all.time, aes(x = percent.dev, y = log2(T1SIZE), colour = interaction(sex,species))) +
  geom_point() +
  geom_vline(aes(xintercept = (0.781)), colour = "red") +
  geom_vline(aes(xintercept = (0.838)), colour = "red") +
  geom_vline(aes(xintercept = (1.2)), colour = "black", linetype = 2) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  xlab(NULL) +
  ylab("T1 Larval Disc size, log2") +
  xlim(c(0.4,1.4)) +
  theme_classic()
```


```{r}
disc.number <- ggplot(all.time, aes(x = percent.dev, y = log2(T1DAPI), colour = interaction(sex,species))) +
  geom_point() +
  geom_vline(aes(xintercept = (0.781)), colour = "red") +
  geom_vline(aes(xintercept = (0.838)), colour = "red") +
  geom_vline(aes(xintercept = (1.2)), colour = "black", linetype = 2) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  xlab("Percent third instar larval development") +
  ylab("T1 Larval cell count, log2") +
  xlim(c(0.4,1.4)) +
  theme_classic()
```


```{r}
disc.size / disc.number + 
  plot_layout(guides = "collect")

```



```{r}
ggplot(all.time, aes(x = percent.dev, y = log2(T2SIZE), colour = interaction(sex,species))) +
  geom_point() +
  geom_vline(aes(xintercept = (0.781)), colour = "red") +
  geom_vline(aes(xintercept = (0.838)), colour = "red") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  xlab("Percent third instar Larval Development") +
  ylab("T2 Larval Disc size, log2") +
  theme_bw()
```


## Gompertz growth model

As an alternative, let's fit an explicit mechanistic growth model. We will only do it for the *D. prolongata* data as we we have sufficient data for it.

```{r}
Dp_dat <- all.time[all.time$species == "D. prolongata",]
Dp_dat <- droplevels(Dp_dat)

Dp_dat2 <- subset(Dp_dat, select = c(T1SIZE, percent.dev, sex))
Dp_dat2 <- na.omit(Dp_dat2)
```

Check the subset
```{r}
ggplot(Dp_dat2, aes(x = percent.dev, y = T1SIZE, colour = sex)) +
  geom_point() +
  geom_vline(aes(xintercept = (0.781)), colour = "red") +
  geom_vline(aes(xintercept = (0.838)), colour = "red") +
  geom_vline(aes(xintercept = (1.2)), colour = "black", linetype = 2) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  xlab(NULL) +
  ylab("T1 Larval Disc size") +
  xlim(c(0.4,1.4)) +
  theme_classic()
```


Gompertz model, getting it working

 For the Gompertz model, here are interpretations of the parameters
growth Asymptote a 
developmental rate constant b
age decay of growth c
add back in time to initiation of growth t0

```{r}
size <- log2(Dp_dat2$T1SIZE)
age <-  (Dp_dat2$percent.dev)
```



```{r}
gompertz.NLL <- function(a1, b1, c1, sigma) {
	size.gomp <- a1*exp(-b1*exp(-c1*age))
	-sum(dnorm(size, mean = size.gomp, sd = sigma, log=T))}
```

```{r}
min(size)
max(size)
min(age)
sd(size)
```

```{r}
gompertz.MLE.1 <- mle2(gompertz.NLL, 
                       start = list(a1 = 15.86, b1 = 7, c1 = 8.7, sigma = 0.75), 
    method = "BFGS")

summary(gompertz.MLE.1)
```



```{r}
gompertz.MLE.a <- mle2(log2(T1SIZE) ~ dnorm(a*exp(-b*exp(-c*percent.dev)), sd = sigma), 
                        start = list(a = 10, b = 5, c = 5, sigma = 0.75), 
                        parameters = list(a ~ sex,b ~ sex, c ~ sex), method="SANN", 
                       data = Dp_dat2)

summary(gompertz.MLE.a)
```

