---
title: "PCA_work"
author: "Tyler Audet"
date: "2024-08-13"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the script for the PCA analyses for the paper 'Genomic architecture of the developing forelegs of Drosophila prolongata; an exaggerated weapon and ornament'. This script reads in raw count data generated from `STAR_[species].sh` and gemerates Figures 2 and 3, as well as supplemental figure S6.

Load packages and counts
```{r}

library(tidyverse)
library(data.table)
library(DESeq2)
library(edgeR)
library(cowplot)
library(patchwork)

```


The following three blocks of code are just to create the data frames needed for the PCA analysis

Load in pro data
```{r}
# Import all the file names
setwd("../../data/pro/")
in_dir = dir(, pattern = "ReadsPerGene.out.tab")
system(paste("wc -l ",in_dir[[1]]))
# read in counts, sort and put together
counts_in <- lapply(in_dir, function(x) read.table(x, header=T, sep = "", nrows = 16695))
counts_sorted <- lapply(counts_in, function(x) x <- x[order(x[,1]),])
tot_count_matrix <- matrix(unlist(lapply(counts_sorted, function(x) x$V2)), 
                           nrow=16695, ncol=36)

colnames(tot_count_matrix) <- in_dir
rownames(tot_count_matrix) <- counts_sorted[[1]][,1]
notGenes <- grep('N_',rownames(tot_count_matrix))
#missingGenes <- grep('Missing',rownames(tot_count_matrix))
tot_count_matrix <- tot_count_matrix[-notGenes,]
#tot_count_matrix <- tot_count_matrix[-missingGenes,]

# writing out as a table so that I can run this on CC
write.table(tot_count_matrix, "../../pro_counts.txt", col.names = T)
```

Load in carrolli
```{r}
# Import all the file names
setwd("../../car/")
in_dir = dir(, pattern="ReadsPerGene.out.tab")
system(paste("wc -l ",in_dir[[1]]))
# read in counts, sort and put together
counts_in <- lapply(in_dir, function(x) read.table(x, header=T, sep = "", nrows = 16331))
counts_sorted <- lapply(counts_in, function(x) x <- x[order(x[,1]),])
tot_count_matrix <- matrix(unlist(lapply(counts_sorted, function(x) x$V2)), 
                           nrow=16331, ncol=37)

colnames(tot_count_matrix) <- in_dir
rownames(tot_count_matrix) <- counts_sorted[[1]][,1]
notGenes <- grep('N_',rownames(tot_count_matrix))
#missingGenes <- grep('Missing',rownames(tot_count_matrix))
tot_count_matrix <- tot_count_matrix[-notGenes,]
#tot_count_matrix <- tot_count_matrix[-missingGenes,]

# writing out as a table so that I can run this on CC
write.table(tot_count_matrix, "../../car_counts.txt", col.names = T)
```

Load in melanogaster data

```{r}
# Import all the file names
setwd("../../mel/")
in_dir = dir(, pattern="ReadsPerGene.out.tab")
system(paste("wc -l ",in_dir[[1]]))
# read in counts, sort and put together
counts_in <- lapply(in_dir, function(x) read.table(x, header=T, sep = "", nrows = 17613))
counts_sorted <- lapply(counts_in, function(x) x <- x[order(x[,1]),])
tot_count_matrix <- matrix(unlist(lapply(counts_sorted, function(x) x$V2)), 
                           nrow=17613, ncol=35)

colnames(tot_count_matrix) <- in_dir
rownames(tot_count_matrix) <- counts_sorted[[1]][,1]
notGenes <- grep('N_',rownames(tot_count_matrix))
#missingGenes <- grep('Missing',rownames(tot_count_matrix))
tot_count_matrix <- tot_count_matrix[-notGenes,]
#tot_count_matrix <- tot_count_matrix[-missingGenes,]

# writing out as a table so that I can run this on CC
write.table(tot_count_matrix, "../../mel_counts.txt", col.names = T)
```


The next three blocks of code are to create the DESeq objects that get loaded into the PCA function


```{r}
pro_counts <- read.table("../../../../outputs/pro_counts.txt", header = T)

# Need to make the experimental design matrix
# This can be taken just from the label I think
pro_names <- as.data.frame(names(as.data.frame(pro_counts)))
colnames(pro_names) <- "sample"

pro_design <- separate(pro_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

pro_design$replicate <- sub('ReadsPerGene.out.tab', "", pro_design$replicate)


pro_design <- pro_design[,-c(1)]

#Need to make my formula for DeSeq

pro_formula <- model.matrix(~ sex * leg * stage + replicate , data = pro_design)

#Naming first column 'gene' and clearing up unnecessary column and first few rows of summary stats
#rownames(mel_counts) <- mel_counts$gene
# <- mel_counts[-c(1:4),-1]


# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(pro_counts))

#Determine which genes have sufficiently large counts to be retained in a statistical analysis.
#keep1 <- filterByExpr(DGE_object, mel_formula)
#DGE_object <- DGE_object[keep, ]

smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I add one to every count to stop complete seperation issues
#DGE_object$counts <- DGE_object$counts + 1

# I want to filter my mel_counts by the DGE_object to get mel_counts down to the 11691 genes with sufficient counts
pro_counts <- pro_counts[row.names(DGE_object$counts),]
#dim(mel_counts)

pro_counts <- pro_counts + 1

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
pro_data.in <- DESeqDataSetFromMatrix(countData = pro_counts,
                                  colData = pro_design,
                                  design = pro_formula)
pro_data.in <- DESeq2::estimateSizeFactors(pro_data.in)
```

```{r}
car_counts <- read.table("../../car_counts.txt", header = T)

#mel_counts <- read.table("../../data/mel_counts.txt", header = T)

#mel_counts <- read.table("~/Desktop/tester.txt", header = T)


# Need to make the experimental design matrix
# This can be taken just from the label I think
car_names <- as.data.frame(names(as.data.frame(car_counts)))
colnames(car_names) <- "sample"

car_design <- separate(car_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

car_design$replicate <- sub('ReadsPerGene.out.tab', "", car_design$replicate)


car_design <- car_design[,-c(1)]

#Need to make my formula for DeSeq

car_formula <- model.matrix(~ sex * leg * stage + replicate , data = car_design)

#Naming first column 'gene' and clearing up unnecessary column and first few rows of summary stats
#rownames(mel_counts) <- mel_counts$gene
# <- mel_counts[-c(1:4),-1]


# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(car_counts))

#Determine which genes have sufficiently large counts to be retained in a statistical analysis.
#keep <- filterByExpr(DGE_object, mel_formula)
#DGE_object <- DGE_object[keep, ]
# goes from 17611 to 11691, so like 600 genes dropped due to low coverage

smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I add one to every count to stop complete seperation issues
#DGE_object$counts <- DGE_object$counts + 1

# I want to filter my mel_counts by the DGE_object to get mel_counts down to the 11691 genes with sufficient counts
car_counts <- car_counts[row.names(DGE_object$counts),]

car_counts <- car_counts + 1

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
car_data.in <- DESeqDataSetFromMatrix(countData = car_counts,
                                  colData = car_design,
                                  design = car_formula)
car_data.in <- DESeq2::estimateSizeFactors(car_data.in)
```


```{r}
mel_counts <- read.table("../../mel_counts.txt", header = T)

#mel_counts <- read.table("../../data/mel_counts.txt", header = T)

#mel_counts <- read.table("~/Desktop/tester.txt", header = T)


# Need to make the experimental design matrix
# This can be taken just from the label I think
mel_names <- as.data.frame(names(as.data.frame(mel_counts)))
colnames(mel_names) <- "sample"

mel_names$sample <- sub('ReadsPerGene.out.tab', "", mel_names$sample)


mel_design <- separate(mel_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)



mel_design <- mel_design[,-c(1)]

#Need to make my formula for DeSeq

mel_formula <- model.matrix(~ sex * leg * stage + replicate , data = mel_design)

#Naming first column 'gene' and clearing up unnecessary column and first few rows of summary stats
#rownames(mel_counts) <- mel_counts$gene
# <- mel_counts[-c(1:4),-1]


# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(mel_counts))

#Determine which genes have sufficiently large counts to be retained in a statistical analysis.
#keep1 <- filterByExpr(DGE_object, mel_formula)
#DGE_object <- DGE_object[keep, ]
# goes from 17611 to 11691, so like 600 genes dropped due to low coverage

smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I add one to every count to stop complete seperation issues
#DGE_object$counts <- DGE_object$counts + 1

# I want to filter my mel_counts by the DGE_object to get mel_counts down to the 11691 genes with sufficient counts
mel_counts <- mel_counts[row.names(DGE_object$counts),]

mel_counts <- mel_counts + 1

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
mel_data.in <- DESeqDataSetFromMatrix(countData = mel_counts,
                                  colData = mel_design,
                                  design = mel_formula)
mel_data.in <- DESeq2::estimateSizeFactors(mel_data.in)
```


Sparsity plots for the three species
```{r}

plotSparsity(pro_data.in, normalized = TRUE, 
             col = rgb(0,0,1, 0.15), pch = 20)
plotSparsity(car_data.in, normalized = TRUE, 
             col = rgb(0,0,1, 0.15), pch = 20)
plotSparsity(mel_data.in, normalized = TRUE, 
             col = rgb(0,0,1, 0.15), pch = 20)

```



```{r}
#vsn and rlog
#for_pca_rlog <- rlog(data.in, 
#                blind = TRUE)


pro_pca <- rlog(pro_data.in, 
                     blind = TRUE)
car_pca <- rlog(car_data.in, 
                     blind = TRUE)
mel_pca <- rlog(mel_data.in, 
                     blind = TRUE)

###

plotPCA(pro_pca, 
        intgroup=c("sex", "leg"),
        ntop = 1000,
        pcsToUse = c(1,2)) +
  theme_bw()

plotPCA(car_pca, 
        intgroup=c("sex", "leg", "stage"),
        ntop = 1000,
        pcsToUse = c(1,2)) +
  theme_bw()

plotPCA(mel_pca, 
        intgroup=c("sex", "leg", "stage"),
        ntop = 1000,
        pcsToUse = c(1,2)) +
  theme_bw()

###

pro_pca_data <- plotPCA(pro_pca, 
        intgroup=c("sex", "leg", "stage"),
        ntop = 1000,
        pcsToUse = c(3,4),
        returnData=TRUE)
pro_percentVar <- round(100 * attr(pro_pca_data, "percentVar"))

car_pca_data <- plotPCA(car_pca, 
        intgroup=c("sex", "leg", "stage"),
        ntop = 1000,
        pcsToUse = c(3,4),
        returnData=TRUE)
car_percentVar <- round(100 * attr(car_pca_data, "percentVar"))

mel_pca_data <- plotPCA(mel_pca, 
        intgroup=c("sex", "leg", "stage"),
        ntop = 1000,
        pcsToUse = c(3,4),
        returnData=TRUE)
mel_percentVar <- round(100 * attr(mel_pca_data, "percentVar"))


colour_vector <- c("for.F" = "#E66100",
                   "mid.F" = "orange",
                   "for.M" = "#5D3A9B",
                   "mid.M" = "orchid")



p1 <- ggplot(pro_pca_data, aes(PC3,PC4, colour=interaction(leg,sex), shape=stage)) +
     geom_point(size=4.5, alpha = 0.5) +
     xlab(paste0("PC3: ",pro_percentVar[1],"% variance")) +
     ylab(paste0("PC4: ",pro_percentVar[2],"% variance")) + 
     coord_fixed() +
  scale_color_manual(values=colour_vector,labels = c("Female foreleg", "Female midleg","Male foreleg", "Male midleg"), name = "Sex and leg") +
  theme_classic() +
  #theme(legend.position="none") +
  labs(title = "D. prolongata") +
  theme(plot.title = element_text(face = "italic")) +
  ylim(c(-26,26)) +
  xlim(c(-26,26)) +
  scale_shape_manual(values = c(16, 1), labels = c("Early development", "Late development"), name = "Stage")


p2 <-  ggplot(car_pca_data, aes(PC3,PC4, colour=interaction(leg,sex), shape=stage)) +
     geom_point(size=4.5, alpha = 0.5) +
     xlab(paste0("PC3: ",car_percentVar[1],"% variance")) +
     ylab(paste0("PC4: ",car_percentVar[2],"% variance")) + 
     coord_fixed() +
  scale_color_manual(values=colour_vector,labels = c("Female foreleg", "Female midleg","Male foreleg", "Male midleg"), name = "Sex and leg") +
  theme_classic() +
  #theme(legend.position="none") +
  labs(title = "D. carrolli") +
  theme(plot.title = element_text(face = "italic")) +
  ylim(c(-26,26)) +
  xlim(c(-26,26)) +
  scale_shape_manual(values = c(15, 0)) +
  theme(legend.position = "none")
#, labels = c("Early development", "Late development"), name = "Stage")


p3 <-  ggplot(mel_pca_data, aes(PC3,PC4, colour=interaction(leg,sex), shape=stage)) +
     geom_point(size=4.5, alpha = 0.5) +
     xlab(paste0("PC3: ",mel_percentVar[1],"% variance")) +
     ylab(paste0("PC4: ",mel_percentVar[2],"% variance")) + 
     coord_fixed() +
   scale_color_manual(values=colour_vector,labels = c("Female foreleg", "Female midleg","Male foreleg", "Male midleg"), name = "Sex and leg") +
  theme_classic() +
  #theme(legend.position="none") +
  labs(title = "D. melanogaster") +
  theme(plot.title = element_text(face = "italic")) +
  ylim(c(-26,26)) +
  xlim(c(-26,26)) +
  scale_shape_manual(values = c(17, 2), labels = c("Early development", "Late development"), name = "Stage") +
  theme(legend.position = "none")


#legend <- get_legend(dummy_PCA_plot)

# patchwork plotting

design <- "AB
           CD"

p1 + p2 + p3 + guide_area() +
  plot_layout(guides = 'collect')

#plot_grid(pro_PCA_plot,car_PCA_plot,mel_PCA_plot,legend)

###




```


I want to put all the species together. First I merge all the overlapping genes.

```{r}

test <- merge(mel_counts, car_counts, by = 0)

rownames(test) <- test$Row.names
test <- test[,-1]

merged_df <- merge(test, pro_counts, by = 0) 
rownames(merged_df) <- merged_df$Row.names

grn <- read.table(file = "../../../../outputs/grn_targetGenes.txt")

merged_grn <- merged_df[merged_df[,1] %in% grn$V1,]

merged_df <- merged_df[,-1]
merged_grn <- merged_grn[,-1]

```


```{r}
merged_names <- as.data.frame(names(as.data.frame(merged_df)))
colnames(merged_names) <- "sample"

merged_design <- separate(merged_names, 
                       col = "sample", 
                       into = c("species","sex","leg","stage", "replicate"),
                       sep = "_", remove = FALSE)

merged_design$replicate <- sub('ReadsPerGene.out.tab', "", merged_design$replicate)


merged_design <- merged_design[,-c(1)]

#Need to make my formula for DeSeq

merged_formula <- model.matrix(~ sex * leg * stage + replicate , data = merged_design)

#Naming first column 'gene' and clearing up unnecessary column and first few rows of summary stats
#rownames(mel_counts) <- mel_counts$gene
# <- mel_counts[-c(1:4),-1]


# create a Dseq differentially expressed gene object
DGE_object <- DGEList(as.matrix(merged_df))

#Determine which genes have sufficiently large counts to be retained in a statistical analysis.
#keep1 <- filterByExpr(DGE_object, mel_formula)
#DGE_object <- DGE_object[keep, ]

smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I add one to every count to stop complete seperation issues
#DGE_object$counts <- DGE_object$counts + 1

# I want to filter my mel_counts by the DGE_object to get mel_counts down to the 11691 genes with sufficient counts
merged_df <- merged_df[row.names(DGE_object$counts),]
#dim(mel_counts)

merged_df <- merged_df + 1

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
merged_data.in <- DESeqDataSetFromMatrix(countData = merged_df,
                                  colData = merged_design,
                                  design = merged_formula)
merged_data.in <- DESeq2::estimateSizeFactors(merged_data.in)

```


```{r}
merged_pca <- rlog(merged_data.in, 
                     blind = TRUE)
```


```{r}
### Everything

plotPCA(merged_pca, pcsToUse = c(1,2),
        intgroup=c("species","sex", "leg", "stage"),
        ntop = 1000) +
  theme_bw()

# for plotting
merged_PCA_data <- plotPCA(merged_pca, pcsToUse = c(1,2),
        intgroup=c("species","sex", "leg", "stage"),
        returnData=TRUE,
        ntop = 1000)
percentVar <- round(100 * attr(merged_PCA_data, "percentVar"))

## Species Figure S3

colour_vector <- c("for.F" = "#E66100",
                   "mid.F" = "orange",
                   "for.M" = "#5D3A9B",
                   "mid.M" = "orchid")

shape_vector <- c("car.early" = 15,
                  "mel.early" = 17,
                  "pro.early" = 16,
                  "car.late" = 0,
                  "mel.late" = 2,
                  "pro.late" = 1)


PC1_PC2 <- ggplot(merged_PCA_data, aes(PC1,PC2, colour=interaction(leg,sex), shape=interaction(species,stage), group = species, )) +
     geom_point(size=4.5, alpha = 0.5) +
     xlab(paste0("PC1: ",percentVar[1],"% variance")) +
     ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
     coord_fixed() +
     scale_color_manual(values=colour_vector,labels = c("Female foreleg", "Male foreleg","Female midleg", "Male midleg"), name = "Sex and leg") +
  theme_classic() +
  scale_shape_manual(values = shape_vector, labels = c(expression(paste(italic("D. carrolli "), "early")), expression(paste(italic("D. melanogaster "), "early")),
expression(paste(italic("D. prolongata "), "early")),
expression(paste(italic("D. carrolli "), "late")),
expression(paste(italic("D. melanogaster "), "late")),
expression(paste(italic("D. prolongata "), "late"))),
name = "Species and stage")

##

merged_PCA_data <- plotPCA(merged_pca, pcsToUse = c(3,4),
        intgroup=c("species","sex", "leg", "stage"),
        returnData=TRUE,
        ntop = 1000)
percentVar <- round(100 * attr(merged_PCA_data, "percentVar"))

## Species Figure S3

PC3_PC4 <- ggplot(merged_PCA_data, aes(PC3,PC4, colour=interaction(leg,sex), shape=interaction(species,stage), group = species, )) +
     geom_point(size=4.5, alpha = 0.5) +
     xlab(paste0("PC3: ",percentVar[1],"% variance")) +
     ylab(paste0("PC4: ",percentVar[2],"% variance")) + 
     coord_fixed() +          scale_color_manual(values=colour_vector,labels = c("Female foreleg", "Male Foreleg","Female midleg", "Male midleg"), name = "Sex and leg") +
  theme_classic() +
  scale_shape_manual(values = shape_vector, labels = c(expression(paste(italic("D. carrolli "), "early")), expression(paste(italic("D. melanogaster "), "early")),
expression(paste(italic("D. prolongata "), "early")),
expression(paste(italic("D. carrolli "), "late")),
expression(paste(italic("D. melanogaster "), "late")),
expression(paste(italic("D. prolongata "), "late"))),
name = "Species and stage") +
  theme(legend.position = "none")
##


(PC1_PC2 / PC3_PC4) +
  plot_layout(guides = 'collect')

```


```{r}

DGE_object <- DGEList(as.matrix(merged_grn))


smallestGroupSize <- 4
keep <- rowSums(DGE_object$counts >= 10) >= smallestGroupSize
DGE_object <- DGE_object[keep,]

# I add one to every count to stop complete seperation issues
#DGE_object$counts <- DGE_object$counts + 1

# I want to filter my mel_counts by the DGE_object to get mel_counts down to the 11691 genes with sufficient counts
merged_grn <- merged_grn[row.names(DGE_object$counts),]
#dim(mel_counts)

merged_grn <- merged_grn + 1

# DESeqDataSet is a subclass of RangedSummarizedExperiment, used to store the input values, intermediate calculations and results of an analysis of differential expression
merged_data.in <- DESeqDataSetFromMatrix(countData = merged_grn,
                                  colData = merged_design,
                                  design = merged_formula)
merged_data.in <- DESeq2::estimateSizeFactors(merged_data.in)


merged_grn_pca <- rlog(merged_data.in, 
                     blind = TRUE)

### Everything

plotPCA(merged_grn_pca, pcsToUse = c(1,2),
        intgroup=c("species","sex", "leg", "stage"),
        ntop = 1000) +
  theme_bw()

# for plotting
merged_PCA_data <- plotPCA(merged_grn_pca, pcsToUse = c(2,3),
        intgroup=c("species","sex", "leg", "stage"),
        returnData=TRUE,
        ntop = 750)
percentVar <- round(100 * attr(merged_PCA_data, "percentVar"))

## Species figure

ggplot(merged_PCA_data, aes(PC2,PC3, colour=interaction(sex,leg), shape=interaction(species,stage), group = species, )) +
     geom_point(size=3, alpha = 0.5) +
     xlab(paste0("PC2: ",percentVar[1],"% variance")) +
     ylab(paste0("PC3: ",percentVar[2],"% variance")) + 
     coord_fixed() +
  #facet_grid(~sex) +
  scale_color_manual(values=c("#E66100","orchid","orange","#5D3A9B")) +
  #stat_ellipse(aes(x=PC1, y=PC2,color=group),type = "norm", alpha = 0.25) +
  theme_classic()
##
```




